---
title: "R Notebook"
output: html_notebook
---
basic.info is the table containing the basic variables for the patient and samples: patientId, sampleid, dates of KT,Event,Sample, rejection status...
basic.info is given as table in the supplement

# Baseline Table 

- table of all baseline variables: Nebgen info, sex, age, Tx info, ... 

```{r}

nebgen.info <- read.csv("../Oct22/NeBGeneral_information.csv", header=T, sep=";" , dec=",")
nebgen.info<- inner_join(nebgen.info, rej.mapping, by=c("PatientID" = "Johannes")) %>% select(-PatientID, -Dokument ) %>% select(PatientID=Daniela, name=X, everything())


baseline.table <- nebgen.info %>% select(-name) %>% distinct() 
## remove constant columns 
baseline.table <- baseline.table %>% remove_empty("cols") %>% remove_constant() 
  
nebgen.info %>% filter(name %in% donor_ids)
  
## remove Date column and allele/HLA columns 
#nebgen.info <- select(nebgen.info, ! matches("Status*|Container*|*Date*|*Datum*|HLA*|allele*")) 
  
baseline.table <- baseline.table %>% select(-Status, -Container.Index, -TxDate.BL.Re.NeBGe) %>% 
                                    rename(Container_Date=Datum..Container.) %>%
                                    mutate(Container_Date= as.Date(Container_Date, format="%d.%m.%Y"), 
                                           Erfassungsdatum= as.Date(Erfassungsdatum, format="%d.%m.%Y"),
                                           InitialdiagDate.BL.Re.NeBGe= as.Date(InitialdiagDate.BL.Re.NeBGe, format="%d.%m.%Y"))

baseline.table <- left_join(basic.info %>% select(PatientID, Transplantation_Date, Event_Date, rej, Geschlecht) %>% distinct(), baseline.table, by="PatientID") %>% relocate(Container, Container_Date)

colnames(baseline.table) <- str_replace(colnames(baseline.table), ".BL.Re.NeBGe", "")
colnames(baseline.table) <- str_replace(colnames(baseline.table), ".BL.Do.NeBGe", "")

baseline.table <- baseline.table %>% mutate(sex=case_when(Geschlecht == "männlich" ~ "male",
                                                          Geschlecht == "weiblich" ~ "female",
                                                          T ~ "unkown"), .keep="unused", .before=height.cm)


X1_NeBGeneral_information_IC_filtered <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Oct22/all_files_containing_patient_information/filtered/1.-NeBGeneral_information_IC_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)


X1_NeBGeneral_information_IC_filtered[81,]$`ICDcode-BL-Re-NeBGe` <- "M31.7"

X1_NeBGeneral_information_IC_filtered<- X1_NeBGeneral_information_IC_filtered %>% select(PatientID, contains("IC") )

names(X1_NeBGeneral_information_IC_filtered) <- str_replace(names(X1_NeBGeneral_information_IC_filtered), "-BL-Re-NeBGe", "")

X1_NeBGeneral_information_IC_filtered <- X1_NeBGeneral_information_IC_filtered %>% select(PatientID,ICDcode, ICDlabel )  %>% group_by(PatientID) %>% mutate(row_id = row_number()) %>% 
  pivot_wider(id_cols = "PatientID",  names_from = row_id,
              values_from = c(ICDcode,ICDlabel) , 
              names_glue = "{.value}{row_id}") %>% 
  ungroup()

baseline.table<- left_join(baseline.table, X1_NeBGeneral_information_IC_filtered , by="PatientID") %>% relocate(starts_with("ICD"), .before=registrationDate)

## error switched height and weight 
# PatientID == "MWD5FRF0"
weight <- baseline.table[59,12]
height <- baseline.table[59,13]
baseline.table[59,12] <- height
baseline.table[59,13] <- weight
baseline.table[59,]
## 21YDKU88
weight <- baseline.table[118,12]
height <- baseline.table[118,13]
baseline.table[118,12] <- height
baseline.table[118,13] <- weight
baseline.table[118,]



## calculate BMI 
baseline.table<- baseline.table %>% mutate(BMI = weightAtTx.kg/((height.cm/100)^2) , BMI.donor=  weightAtTx.donor.kg/((height.donor.cm/100)^2),.before=height.cm)




Personenvariablen <- read_delim("Oct22/all_files_containing_patient_information/Personenvariablen.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

baseline.table <- baseline.table %>% left_join(Personenvariablen %>% select(PatientID, Site), by="PatientID") %>% relocate(Site, .before=sex)

baseline.table %>% filter((!PatientID %in% out_patients) & PatientID %in% cohort2_patients) %>% pull(Site) %>% table()/92
baseline.table %>% filter((!PatientID %in% out_patients) & PatientID %in% cohort2_patients) %>% pull(Site) %>% table()
baseline.table %>% filter((!PatientID %in% out_patients) & PatientID %in% cohort2_patients) %>% count(Site,rej) %>% mutate(perc=case_when(rej==0 ~ n/60, T ~ n/32))
baseline.table %>% filter((!PatientID %in% out_patients) & PatientID %in% cohort2_patients) %>% count(rej,Site) %>% group_by(Site) %>% mutate(perc=n/sum(n))

baseline.table  %>% pull(Site) %>% table()/231


## blood type 

baseline.table<- baseline.table %>% mutate(bloodAB0.same= case_when(bloodAB0 == bloodAB0.donor ~ "yes", T ~ "no"), rhesus.same= case_when(bloodRhesus == bloodRhesus.donor ~ "yes", T ~ "no"), .before=bloodAB0)


```



# Labortaory Values 

## Reading Files 
- 1.2 Laboratory file : "Datum..Container..x.x"  , BL.Re.NeBLa 
- 2.2 old Laboratory file: FU-Re-NeFLa
- LabBio folder: Laboratory.csv BL-Re-Labor with also Donor blood values 

### Lab File 1.2

```{r}
X1.2.Lab <- mergeContainer %>% select(Container, Container_Date=Datum..Container..x.x, PatientID, ends_with("BL.Re.NeBLa") ) %>% distinct()  %>% mutate(across(everything(), as.character))

colnames(X1.2.Lab) <- str_replace(colnames(X1.2.Lab) , ".BL.Re.NeBLa", "")

X1.2.Lab<- X1.2.Lab  %>% filter(rowSums(is.na(X1.2.Lab)) < ncol(X1.2.Lab)-3) 


X2_2Laboratory_filtered <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Oct22/all_files_containing_patient_information/filtered/2.2Laboratory_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)  %>% mutate(across(everything(), as.character))

colnames(X2_2Laboratory_filtered) <- str_replace(colnames(X2_2Laboratory_filtered), "-FU-Re-NeFLa", "")
colnames(X2_2Laboratory_filtered) <- str_replace_all(colnames(X2_2Laboratory_filtered), "-", ".")

## filter out empty rows , KEIN DOKUMENT
X2_2Laboratory_filtered <- X2_2Laboratory_filtered %>% filter(Dokument!= "KEIN DOKUMENT")

X2_2Laboratory_filtered <- X2_2Laboratory_filtered %>% select(-Dokument, -Erfassungsdatum, -`Container Index`, -Status, -contains("AGEF")) %>%
                                                       rename(Container_Date= `Datum (Container)`)


X2_2Laboratory_filtered


```


### LabBio file: 

- with donor info , separating it and save it an extra file 


```{r}

## read in, changing col names 
Laboratory_filtered <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Oct22/all_files_containing_patient_information/filtered/Laboratory_filtered.csv", delim = ";", escape_double = FALSE, trim_ws = TRUE)  %>% mutate(across(everything(), as.character))

colnames(Laboratory_filtered) <- str_replace(colnames(Laboratory_filtered), "-BL-Re-Labor", "")
colnames(Laboratory_filtered) <- str_replace_all(colnames(Laboratory_filtered), "-", ".")


###############################################
#extracting donor information 

Donor_Lab <- Laboratory_filtered %>% select(Container, Container_Date=`Datum (Container)`, PatientID, Cause,dateOfAscertainment,analyzedWithinTimeRange,dateOfAscertainmentEnd, ends_with("BL.Do.Labor") ) %>% filter(DonLabAvailable.BL.Do.Labor == "yes")


### adding basic info about patients 

Donor_Lab <- right_join(basic.info %>% select(PatientID, Transplantation_Date, Event_Date, rej) %>% distinct() , Donor_Lab, by="PatientID")

### make it more pretty, changing it to Date and relocate Container 

Donor_Lab <- Donor_Lab %>% relocate(Container,Container_Date) %>% 
                           mutate(Container_Date= as.Date(Container_Date, format="%d.%m.%Y")) %>%
                           mutate(Ascertainment_Date= as.Date(dateOfAscertainment, format="%d.%m.%Y"),
                                  Ascertainment_End_Date= as.Date(dateOfAscertainmentEnd, format="%d.%m.%Y"),
                                  .keep="unused", .before=Cause
                                  )

write.table(Donor_Lab %>% arrange(PatientID,Ascertainment_Date),
            "../tables/match_group/Donor_Laboratory.tsv",  quote = F, col.names = T, row.names = F, sep="\t", na="" )                              


########################################

### Recipient 

Laboratory_filtered <- Laboratory_filtered %>% select(-Dokument, -Erfassungsdatum, -`Container Index`, -Status, -contains("AGEF")) %>%
                                                       rename(Container_Date= `Datum (Container)`) %>%
                                              select(-ends_with("BL.Do.Labor"))
colnames(Laboratory_filtered)
```


## Merging Files 

```{r}


all(intersect(colnames(X1.2.Lab), colnames(Laboratory_filtered)) %in% colnames(X1.2.Lab))
all(intersect(colnames(X2_2Laboratory_filtered), colnames(Laboratory_filtered)) %in% colnames(X2_2Laboratory_filtered))

## join them 
Laboratory<- full_join(full_join(X1.2.Lab , X2_2Laboratory_filtered, by=colnames(X1.2.Lab)), Laboratory_filtered, by=colnames(X2_2Laboratory_filtered))

## sort alphabetically
Laboratory<- Laboratory[, order(names(Laboratory))]

## add basic info 
Laboratory <- basic.info %>% select(PatientID, Transplantation_Date, Event_Date, rej) %>% distinct() %>% right_join(Laboratory, by="PatientID")

## relocate columns and change to dates 

Laboratory <- Laboratory %>% relocate(Container, Container_Date) %>% 
                             mutate(Container_Date= as.Date(Container_Date, format="%d.%m.%Y")) %>%
                             mutate(Ascertainment_Date= as.Date(dateOfAscertainment, format="%d.%m.%Y"),
                                  Ascertainment_End_Date= as.Date(dateOfAscertainmentEnd, format="%d.%m.%Y"),
                                  .keep="unused", .after=rej
                                  ) %>% relocate(Cause, analyzedWithinTimeRange, .after=Ascertainment_End_Date)


Laboratory <- as.data.frame(apply(Laboratory, 2, gsub, patt=",", replace="."))


Laboratory<- Laboratory %>% filter(PatientID %in% patientIDs)


```

## Lab Subset 

just interest in a subset of the ascertained values 

```{r}
subset_lab <- Laboratory %>% select(contains("Container") , contains("Date") , PatientID, rej, Cause, analyzedWithinTimeRange, 
                         contains("Albumin"),
                         contains("CRP"),
                         contains("Creatinine"),
                         contains("HbA1c"),
                         contains("Hemoglobin"),
                         contains("Leukocytes"),
                         contains("Lymphocytes"),
                         contains("Monocytes"), 
                         contains("neutro"),
                         contains("Phosphate"),
                         contains("PTH"),
                         contains("Urea"),
                         contains("Uricacid"),
                         contains("Urine.Erythrocytes"),
                         contains("Urine.Leukocytes"), 
                         contains("Urine.Nitrite"), 
                         contains("Urine.Protein"))


subset_lab[,11:ncol(subset_lab)]<- subset_lab[,11:ncol(subset_lab)] %>% mutate(across(-c(ends_with("Unit"), Urine.Nitrite), as.numeric))

subset_lab <- subset_lab %>% mutate_at(vars(ends_with("_Date")), as.Date, format="%Y-%m-%d")
subset_lab <- subset_lab %>% mutate(Ascertainment_Tx_diffs = Ascertainment_Date - Transplantation_Date, 
                                    Ascertainment_Event_diffs = Ascertainment_Date - Event_Date, .before=PatientID )


lab.tables <- apply(subset_lab[,11:ncol(subset_lab)]  %>% select(is.numeric) ,2, summary)
```

## Value Correction 
### Albumin 
```{r}
## Albumin 
## between 3.5 and 5.5 , 35-55

subset_lab %>% select(contains("Albumin")) %>% remove_empty("rows")

subset_lab <- subset_lab %>% mutate(AlbuminAbs= case_when(
  is.na(AlbuminAbs) & !is.na(AlbuminRel) ~ AlbuminRel ,
  T ~ AlbuminAbs)) %>% 
  mutate(Albumin= case_when(
  AlbuminAbs > 100 ~ AlbuminAbs/10,
  AlbuminAbs < 10 ~ AlbuminAbs*10,
  T ~ AlbuminAbs
)) %>% select(-AlbuminAbs, -AlbuminRel)

boxplot(subset_lab$Albumin)
```

### CRP 
```{r}
##  CRP: CRPfinal is converted already in mg/L, 
#180 values were lost due to missing unit or sometimes just not added. 
## checked the cause for taking the lab, no rejection/infection cause -> treating it as already changed to mg/L thus lower CRP values -> normal measures

subset_lab <- subset_lab %>%
  mutate(CRP_critical= case_when(!is.na(CRP) & is.na(CRPFinal) & is.na(CRPUnit) ~ "yes",
                               T ~ "no"),
           CRPFinal = case_when(!is.na(CRPFinal) ~ CRPFinal,
                              !is.na(CRP) & is.na(CRPFinal) & CRPUnit == "mg/l" ~ CRP,
                              !is.na(CRP) & is.na(CRPFinal) & CRPUnit == "mg/dl" ~ CRP*10,
                              ## critical move
                              !is.na(CRP) & is.na(CRPFinal) & is.na(CRPUnit)  ~ CRP,
                              is.na(CRP) & is.na(CRPFinal) ~ NA
                              
                              ), .after=CRP) %>% select(-CRP, -CRPUnit)
```

### Creatinine
```{r}
## Creatinine : normal< 1.3, one kidney: 1.8-1.9 , above abnormal > 5.0 severe kidney damage
## more than 200 creatinine values but no final and unit , taken the value as they are, since only mg/dl is reported , labeled in Creatinine_Critical as yes 

subset_lab<- subset_lab %>% mutate(Creatinine_Critical = case_when(is.na(CreatinineFinal) & !is.na(Creatinine) & is.na(CreatinineUnit) ~ "yes", T ~ "no"), .after=Creatinine) %>%
              mutate (CreatinineFinal= case_when(
                        # critical
                       is.na(CreatinineFinal) & !is.na(Creatinine) ~ Creatinine,
                       T ~ CreatinineFinal ), .after=Creatinine) %>% 
              select(-Creatinine, -CreatinineUnit)
```

### Hemoglobin
```{r}
## Hemoglobin, 
#no given unit, chaging it into g/dl 

boxplot(subset_lab$Hemoglobin)

## critical changes: 
## everything smaller than 50 -> *10
### everything bigger than 1000 -> /10

## goal: unit of g/L normal range 138-172 g/L for men 

subset_lab<- subset_lab %>% mutate(HemoglobinFinal = case_when(
  Hemoglobin <= 50 ~ Hemoglobin*10,
  Hemoglobin >= 1000 ~ Hemoglobin/10,
  T ~ Hemoglobin
), .after=Hemoglobin) %>% mutate(
  Hemoglobin_Critical = case_when(Hemoglobin != HemoglobinFinal ~ "yes", T ~ "no"),.after=Hemoglobin
) 

ggplot(select(subset_lab, before=Hemoglobin, after=HemoglobinFinal) %>% gather("type", "value") , aes(x=type,y=value) ) + geom_boxplot()       


summary(subset_lab$HemoglobinFinal)

subset_lab <- select(subset_lab, -Hemoglobin)
```

### Leukocytes
```{r}
### Leukocytes: 4500 - 9000 normal 

# goal: changing it into 1/µl 
## mistake by conversion of 1/nl, should be the same way converted as G/l 

ggplot(select(subset_lab,bef=Leukocytes, after=LeukocytesFinal) %>% gather("type", "value"), aes(x=type, y=value)) + geom_boxplot() + ylim(0,50)

summary(subset_lab$Leukocytes)

summary(subset_lab$LeukocytesFinal)

subset_lab<- subset_lab %>%
   mutate(Leukocytes_new= case_when(
     is.na(LeukocytesFinal) & !is.na(Leukocytes) & (LeukocytesUnit=="G/l"|LeukocytesUnit=="1/nl") ~ Leukocytes*1000,
     ## error measure
     (Leukocytes == 999000 | Leukocytes== 0.01027) ~ NA, 
     ## critical moves 
     ## low ones, assume to be in G/l 
     is.na(LeukocytesFinal) & is.na(LeukocytesUnit) & Leukocytes <= 20 ~ Leukocytes*1000,
     ## weird high ones 
     is.na(LeukocytesFinal) & is.na(LeukocytesUnit) & Leukocytes > 50000 ~ Leukocytes/10,
     is.na(LeukocytesFinal) & is.na(LeukocytesUnit) & Leukocytes > 100000 ~ Leukocytes/100,
     
     ## in the accepted range
     is.na(LeukocytesFinal) & Leukocytes >= 500 & Leukocytes <= 50000 ~ Leukocytes,
     
     ## for the already incorrectly converted units 
     !is.na(LeukocytesFinal) & LeukocytesUnit == "1/nl" ~ Leukocytes*1000,
     ## sometimes 1/ul but clearly to low values
     !is.na(LeukocytesFinal) & LeukocytesFinal <= 20 ~ Leukocytes*1000,
     T ~ LeukocytesFinal
  )
           ) %>% mutate(Leukocytes_Critical = case_when(Leukocytes_new != LeukocytesFinal ~ "yes",Leukocytes_new == LeukocytesFinal ~ "no" ))

ggplot(select(subset_lab,new=Leukocytes_new, before=LeukocytesFinal) %>% gather("type", "value"), aes(x=type, y=(value))) + geom_boxplot() + ylim(500,20000)

## changing missing one by hand
subset_lab[929,]$Leukocytes_new <- 8390


summary(subset_lab$Leukocytes_new)
summary(subset_lab$LeukocytesFinal)


## showing best the results: minimizing numbers of outliers 
subset_lab %>% filter(Leukocytes_new > 20000 | Leukocytes_new < 500)

subset_lab %>% filter(LeukocytesFinal > 20000 | LeukocytesFinal < 500)


subset_lab <- subset_lab %>% select(-Leukocytes, -LeukocytesFinal, -LeukocytesUnit) %>% relocate(Leukocytes_Critical,LeukocytesFinal = Leukocytes_new, .after=HemoglobinFinal)


```


### Lymphocytes 
```{r}
# Lymphocytes 

# normal: 1,000 and 4,800 lymphocytes in every 1 microliter 


summary(subset_lab$LymphocytesAbsFinal)
subset_lab<- subset_lab %>% mutate(LymphocytesAbsFinal2= case_when(
                              LymphocytesAbs < 5 ~ LymphocytesAbs * 1000,
                              T ~ LymphocytesAbs), 
                              .after=LymphocytesAbsFinal)
 
ggplot(subset_lab %>%select(starts_with("LymphocytesAbs") & ! contains("Unit") ) %>% gather(type, value), aes(x=type, y=as.numeric(value))) + geom_boxplot()
```

### monocytes
```{r}
summary(subset_lab$MonocytesAbs)
summary(subset_lab$MonocytesAbsFinal)

subset_lab<- subset_lab %>% mutate(MonocytesAbsFinal2= case_when(
                              MonocytesAbs > 100 ~ MonocytesAbs / 1000,
                              T ~MonocytesAbs), 
                              .after=MonocytesAbsFinal)

summary(subset_lab$MonocytesAbsFinal2)

ggplot(subset_lab %>%select(starts_with("MonocytesAbs") & ! contains("Unit") ) %>% gather(type, value), aes(x=type, y=as.numeric(value))) + geom_boxplot() + ylim(0,2000)

ggplot(subset_lab %>%select(MonocytesAbsFinal2,MonocytesAbs )  %>% gather(type, value), aes(x=type, y=as.numeric(value))) + geom_boxplot() + ylim(0,2000)
```

### Total neutrophilis  

```{r}
summary(subset_lab$TotalneutroAbs)
summary(subset_lab$TotalneutroAbsFinal)

subset_lab %>% select(starts_with("TotalneutroAbs")) %>% remove_empty("rows")
  
  
subset_lab <- subset_lab %>%  mutate( TotalneutroAbsFinal2= case_when(
                              TotalneutroAbs >= 1000 ~ TotalneutroAbs / 1000,
                              T ~  TotalneutroAbs), 
                              .after= TotalneutroAbsFinal)
 
ggplot(subset_lab %>%select(starts_with("TotalneutroAbs") & ! contains("Unit") ) %>% gather(type, value), aes(x=type, y=as.numeric(value))) + geom_boxplot()
```

### Phosphate 

```{r}
## nothing to change here , also difficult conversion mmol/l to mg/dl 
summary(subset_lab$PhosphateFinal)
summary(subset_lab$Phosphate)
```

### Urea 

```{r}

## some values are randomly taken *100 despite same unit 
# taken urea instead of ureafinal

subset_lab %>% select(starts_with("Urea"))

summary(subset_lab$UreaFinal)
summary(subset_lab$Urea)
```

### Urine Values
```{r}
## Urine Protein , nothing to change : range of 0 - 20 g protein per liter, whereas < 0.3g is normal and > 20g is very high 
subset_lab %>% select(contains("Protein")) %>% remove_empty("rows")

boxplot(subset_lab$Urine.ProteinFinal)

## Urine.Erythrocytes, also nthing to change, only 57 entries 

subset_lab %>% select(contains("Urine.Leukocytes")) %>% remove_empty("rows")

```

## Matching Laboratory to Stool Samples

```{r}
match.Laboratory <- subset_lab %>% full_join(basic.info %>% select(SampleID, PatientID, SampleDate=Date),., by="PatientID")




match.Laboratory <- match.Laboratory %>% filter(PatientID %in% patientIDs & !is.na(SampleDate)) %>% 
  mutate( distance= SampleDate - as.Date(Ascertainment_Date, format="%Y-%m-%d"), .before=Container )%>% relocate(Ascertainment_Date, .after=SampleDate) %>%
  group_by(PatientID, SampleID) %>% slice_min(abs(distance)) %>% select(-analyzedWithinTimeRange) %>% distinct() %>% ungroup()

match.Laboratory$SampleID[table(match.Laboratory$SampleID) > 1]


match.Laboratory %>% group_by(SampleID) %>% summarise(n=n()) %>% filter(n>1)

match.Laboratory<- match.Laboratory%>% group_by(SampleID) %>% slice_max(UreaFinal)

match.Laboratory<- match.Laboratory %>% rename(Lab_Sample_Distance=distance)
## 3 samples without lab values since they also not have any sample date 
```



## Creatinine Timelines & Plots 

### Timelines
#### All Follow Ups
```{r}
CreatinineTable <- subset_lab %>% select(PatientID, Ascertainment_Date,  Ascertainment_Tx_diffs, Ascertainment_Event_diffs, rej, starts_with("Crea")) %>% filter(!is.na(CreatinineFinal)) %>% arrange(rej, PatientID) %>% filter(PatientID %in% cohort2_patients)

creatinine_plotdf <- CreatinineTable %>% filter(PatientID %in% cohort2_patients & CreatinineFinal < 30 ) %>% mutate(type=case_when(Ascertainment_Event_diffs ==0 ~ "on rejection day", Ascertainment_Tx_diffs ==0 ~ "on KT day", T~ "not"))

coloring <- c("on rejection day" = "#0072B5FF","on KT day" = "#BC3C29FF","not"="black" )

    ggplot(creatinine_plotdf %>% mutate(rej=case_when(rej==0 ~ "Normal Progress", T ~ "Rejection")), 
         aes(x=as.numeric(Ascertainment_Tx_diffs), y=(CreatinineFinal))) + 
        facet_wrap( ~rej,scales="free")+  geom_line(aes(group=PatientID), linetype="dashed") + 
          geom_smooth() + geom_point(aes(color=type), size = 2, shape = 16) + scale_y_continuous(limits = c(0, 10)) + xlab("Day Difference of Ascertainment and KT") + ylab("Creatinine in mg/dL")+
              theme(legend.position = "top") + scale_color_manual(values = coloring,breaks = c("on rejection day" ,"on KT day")) +  theme_bw() + 
   theme(
    strip.text = element_text(size=16),
    strip.background = element_rect(fill="white", linewidth = 0),
    line = element_line(colour = "darkgrey"),
    axis.ticks.y  = element_line(color = "black"),
    axis.text.y = element_text(color="black", size=11),
    
    axis.text.x = element_text(color="black", size=11),
    axis.ticks.x = element_line(color = "black"),
    
    axis.title.x  =element_text(size=13),
    axis.title.y  =element_text(size=13),

    legend.title = element_blank(),
    legend.position="right",
    legend.text = element_text(color="black", size=11),
    
   
    panel.background = element_blank())
```




```{r}
ggplot(creatinine_plotdf %>% filter(rej==0), 
         aes(x=as.numeric(Ascertainment_Tx_diffs), y=CreatinineFinal)) + 
            geom_line(aes(group=PatientID)) + 
             geom_point(aes(color=type), size = 3, shape = 16) + 
              theme(legend.position = "none") + scale_color_manual(values = coloring) +
              #geom_smooth() + 
              xlab("Day Distance to Tx")


ggplot(creatinine_plotdf %>% filter(type != "on Tx day"), aes(x=interaction(type, rej), y=CreatinineFinal)) + geom_boxplot()


## non rejector with high creatinine values 
weird_nonrej_crea <- CreatinineTable%>% group_by(PatientID) %>% filter(any(CreatinineFinal >= 2.5 & Ascertainment_Tx_diffs > 0) & PatientID %in% cohort2_patients & rej==0) %>% select(-rej,-Ascertainment_Event_diffs)

```


```{r}
## sample based 

CreatinineTable %>% group_by(PatientID) %>% filter(Ascertainment_Tx_diffs >= 0 ) %>%  slice_min(CreatinineFinal) %>% distinct(rej, PatientID, CreatinineFinal) %>% mutate(rej= case_when(rej == 0 ~ "Normal Progress", T ~ "Rejection")) %>% ggplot( aes(x=rej, y=CreatinineFinal,fill=rej)) + geom_boxplot() + stat_compare_means(size=5,label.x.npc = "center") + ylab("Minimum Creatinine") + xlab("")  + theme(legend.position = "none")


CreatinineTable %>% group_by(PatientID) %>% filter(Ascertainment_Tx_diffs > 0) %>% mutate(rej= case_when(rej == 0 ~ "Normal Progress", T ~ "Rejection")) %>% filter(CreatinineFinal <= 20)  %>% ggplot( aes(x=rej, y=CreatinineFinal, fill=rej)) + geom_boxplot() + stat_compare_means(size=5,label.x.npc = "left") + ylab("Creatinine") + xlab("")  + theme(legend.position = "none")


cor.test(CreatinineTable %>% filter(rej==1) %>% filter(Ascertainment_Event_diffs <=0 & Ascertainment_Tx_diffs > 0) %>% pull(CreatinineFinal),
    CreatinineTable %>% filter(rej==1) %>% filter(Ascertainment_Event_diffs <=0& Ascertainment_Tx_diffs > 0)%>% pull(Ascertainment_Event_diffs) %>% as.numeric(),
    method=c( "pearson"))

CreatinineTable %>% filter(rej==1) %>% filter(Ascertainment_Event_diffs <=0 & Ascertainment_Tx_diffs > 0) %>% ggplot(aes(x=Ascertainment_Event_diffs, y=CreatinineFinal)) + geom_point()
```

### Cohort2 Plots
#### Minimum Crea before Sample
```{r}
MinCreatinine_cohort2 <- CreatinineTable %>% filter(Ascertainment_Tx_diffs > 0) %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% cohort2_Ids), by="PatientID") %>% mutate(Ascertainment_Sample_diffs = Ascertainment_Date - SampleDate ) %>% filter(Ascertainment_Sample_diffs <= 0) %>% group_by(SampleID) %>%  slice_min(CreatinineFinal) 


MinCreatinine_cohort2 %>% filter(SampleID %in% in_samples & SampleID != "T024769_S169") %>% distinct(SampleID, CreatinineFinal,rej) %>% mutate(rej= case_when(rej == 0 ~ "Normal Progress", T ~ "Rejection"))%>% ggplot( aes(x=rej, y=CreatinineFinal, fill=rej)) + geom_boxplot() + stat_compare_means(size=5,label.x.npc = "center") + ylab("Minimum Creatinine") + xlab("")  + theme(legend.position = "none") + ylim(0,9)

MinCreatinine_cohort2 %>% filter(SampleID %in% in_samples & SampleID != "T024769_S169") %>% distinct(SampleID, CreatinineFinal,rej) %>% mutate(rej= case_when(rej == 0 ~ "Normal Progress", T ~ "Rejection"))%>% ggplot( aes(x=rej, y=CreatinineFinal, fill=rej)) + geom_boxplot() + stat_compare_means(size=5,label.x.npc = "center") + ylab("Minimum Creatinine") + xlab("")  + theme(legend.position = "none") 

 CreatinineTable %>% filter(Ascertainment_Tx_diffs > 0) %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% cohort2_Ids), by="PatientID") %>% mutate(Ascertainment_Sample_diffs = Ascertainment_Date - SampleDate ) %>% filter(Ascertainment_Sample_diffs > 0)  %>% filter(SampleID %in% in_samples) %>% group_by(SampleID) %>%  slice_min(CreatinineFinal) %>% distinct(PatientID, CreatinineFinal,rej) %>%
 ggplot( aes(x=rej, y=CreatinineFinal, fill=rej)) + geom_boxplot() + stat_compare_means(size=5,label.x.npc = "center") + ylab("Minimum Creatinine") + xlab("")  + theme(legend.position = "none") 

```
#### Last Available Crea

```{r}
LastAvailCreas<- CreatinineTable %>% filter(Ascertainment_Tx_diffs > 0 & PatientID %in% in_patients) %>% distinct() %>% group_by(PatientID) %>% slice_max(Ascertainment_Tx_diffs)


lastavail_crea_plot<- ggplot(LastAvailCreas, aes(x=rej, fill=rej, y=CreatinineFinal)) + geom_boxplot()  + stat_compare_means(size=4,label.x.npc = "center", hjust=0.5) + ylab("Last Available Creatinine after KT") + xlab("")  + 
  scale_x_discrete(labels=c("0"="Non Rejection", "1"="Rejection"))+
  scale_fill_manual(na.translate = F,  values=c( "#0072B5FF","#BC3C29FF")) + theme_classic()+
  theme(
   # axis.line = element_line(colour = "darkgrey"),
    axis.ticks.y  = element_line(color = "black"),
    axis.text.y = element_text(color="black", size=11),
    
    axis.text.x = element_text(color="black", size=11),
    axis.ticks.x = element_line(color = "black"),
    
    axis.title.x  =element_text(size=11),
    axis.title.y  =element_text(size=11),

    legend.title = element_blank(),
    legend.text =  element_text(color="black", size=11),
    legend.position = "none",

    panel.border = element_blank(),
    panel.background = element_blank()
    ) 

LastAvailCreas %>% ungroup() %>% group_by(rej) %>% summarise(mean=mean(Ascertainment_Tx_diffs), median=median(Ascertainment_Tx_diffs), sd=sd(Ascertainment_Tx_diffs))

ggsave("../plots/Figure3/C2_LastAvailableCrea.pdf", width = 5.5, height=3.5)

```



## Creating Variables for Crea
```{r}
allCreatinine_c2 <- CreatinineTable %>% full_join(basic.info %>% filter(SampleID %in% cohort2_Ids) %>% select(SampleID, PatientID, Sample_Date=Date),., by="PatientID") %>%  mutate(Ascertainment_Sample_diffs=  Ascertainment_Date- Sample_Date, .after=Ascertainment_Date) %>% select(-PatientID,-Ascertainment_Date, -Sample_Date,-rej)%>%
  group_by(SampleID) %>% mutate(row = row_number()) %>% rename(Creatinine=CreatinineFinal) %>%
   tidyr::pivot_wider(names_from = row, values_from = c( Ascertainment_Sample_diffs, Ascertainment_Tx_diffs, Ascertainment_Event_diffs,  Creatinine,Creatinine_Critical))
```


```{r}
mean_creatinine_c2<- CreatinineTable %>% full_join(basic.info %>% filter(SampleID %in% cohort2_Ids) %>% select(SampleID, PatientID, Sample_Date=Date),., by="PatientID") %>%  mutate(Ascertainment_Sample_diffs=  Ascertainment_Date- Sample_Date, .after=Ascertainment_Date) %>% group_by(SampleID,rej) %>% summarise(
  mean_total_crea= mean(CreatinineFinal[Ascertainment_Tx_diffs >5], na.rm=T),
  
  mean_before_sample_crea= mean(CreatinineFinal[Ascertainment_Sample_diffs <=0 & Ascertainment_Tx_diffs >0], na.rm=T),
  min_before_sample_crea= min(CreatinineFinal[Ascertainment_Sample_diffs <=0 & Ascertainment_Tx_diffs >0], na.rm=T),
  max_before_sample_crea= max(CreatinineFinal[Ascertainment_Sample_diffs <=0 & Ascertainment_Tx_diffs >0], na.rm=T),

  mean_after_sample_crea= mean(CreatinineFinal[Ascertainment_Sample_diffs >0 & Ascertainment_Tx_diffs >0], na.rm=T),
  min_after_sample_crea= min(CreatinineFinal[Ascertainment_Sample_diffs >0 & Ascertainment_Tx_diffs >0], na.rm=T),
  max_after_sample_crea= max(CreatinineFinal[Ascertainment_Sample_diffs >0 & Ascertainment_Tx_diffs >0], na.rm=T),

  
  mean_before_rej_crea=mean(CreatinineFinal[Ascertainment_Event_diffs < 0 & Ascertainment_Tx_diffs >0], na.rm=T),
  min_before_rej_crea= min(CreatinineFinal[Ascertainment_Event_diffs < 0 & Ascertainment_Tx_diffs >0], na.rm=T),
  max_before_rej_crea= max(CreatinineFinal[Ascertainment_Event_diffs < 0 & Ascertainment_Tx_diffs >0], na.rm=T),

  mean_after_rej_crea= mean(CreatinineFinal[Ascertainment_Event_diffs >=0 & Ascertainment_Tx_diffs >0], na.rm=T),
  min_after_rej_crea= min(CreatinineFinal[Ascertainment_Event_diffs >=0 & Ascertainment_Tx_diffs >0], na.rm=T),
  max_after_rej_crea= max(CreatinineFinal[Ascertainment_Event_diffs >=0 & Ascertainment_Tx_diffs >0], na.rm=T),

)


mean_creatinine_c2[sapply(mean_creatinine_c2, is.nan )] <- NA
mean_creatinine_c2[sapply(mean_creatinine_c2, is.infinite )] <- NA

  
mean_creatinine_c2 %>% select(SampleID, rej, ends_with("sample_crea"))  %>% gather("bef_aft", "val", mean_before_sample_crea, mean_after_sample_crea ) %>% ggplot(aes(y=val, fill=bef_aft, x=rej)) + geom_boxplot() + ylim(0,6) + stat_compare_means() 

mean_creatinine_c2 %>% filter(rej ==1)  %>% gather("bef_aft", "val", mean_before_rej_crea, mean_after_rej_crea ) %>% ggplot(aes(y=val, x=bef_aft, fill=rej)) + geom_boxplot() + ylim(0,6) + stat_compare_means() 

```



