---
title: "Prophylaxe and Infections"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dpylr)
library(tidyr)
library(ggplot2)
```
basic.info is the table containing the basic variables for the patient and samples: patientId, sampleid, dates of KT,Event,Sample, rejection status...
basic.info is given as table in the supplement

#0. Reading Filtered Tables 

those are all files merged by either identical container, if this information is given in the file or merged by PatientID if not
```{r}
mergeContainer <- read.csv("../Oct22/all_files_containing_patient_information/filtered/mergeContainer.tab", sep="\t", header=T)
mergePatient <- read.csv("../Oct22/all_files_containing_patient_information/filtered/mergePatientID.tab", sep="\t", header=T)

colnames(mergeContainer)
```

## General Checks 

some check up of those created files
G945MJK9 with 108 rows !
```{r}
G945MJK9_df <- mergeContainer %>% filter(PatientID == "G945MJK9")
#distinct(as.data.frame(t( G945MJK9_df)))

#distinct(as.data.frame(t(G945MJK9_df$Container)))

filtered.col <- t(as.data.frame(t( G945MJK9_df)) %>% rowwise() %>%  filter(!all(c_across(everything()) == first(c_across(everything())))) %>% ungroup())

## in fact all different entries due to an infection and different treatments 
(as.data.frame(filtered.col) %>% distinct())

```


any intersection container and sample container ? -> no 
```{r}
basic.info %>% select(PatientID, SampleID=name, Container) %>% filter(Container %in%  mergeContainer$Container)
basic.info %>% select(PatientID, SampleID=name, Container) %>% filter(PatientID %in%  mergeContainer$PatientID)

## "intresting" dates if matching with continer is not working 
mergeContainer %>% select(contains("date") & !contains("AGE")) %>% colnames()
```


different files for prophylaxis, antimicrobil treatments, infections and complications 

1.3.4 Prophylaxis 
1.3.4 AntiBacterial 
1.3.4 NeBhbvST and NeBhcvST only one entry each -> ignoring them 
1.3.4 AntiViral 
1.3.4 AntiFungal 
1.3.4 Anti-infectiveprophylax -> summary of the files above but capturing the BL values , no container , PatientID based 

3.1.4 Prophylax files , either empty or almost empty with only one row of information -> ignoring them 

missing end of adminstriation  or duration info 

#1. Prophylaxis

column names containging "Pr" are refering to Prophylaxis

```{r}
# Bacterial Prophylax 163 patients 
mergeContainer %>% filter(!is.na(AntibacName.NeBPr.NeBAn))  %>% filter(PatientID %in% patientIDs) %>% summarise(n=n_distinct(PatientID))

### Viral Prophylax 135 patients
mergeContainer %>% filter(!is.na(AntiviralName.NeBPr.NeBAn)) %>% filter(PatientID %in% patientIDs) %>% summarise(n=n_distinct(PatientID))

### Fungal prophylax 115 patients
mergeContainer %>% filter(!is.na(AntifungName.NeBPr.NeBAn)) %>% filter(PatientID %in% patientIDs) %>% summarise(n=n_distinct(PatientID))

```
##1.1 Bacterial Prophylax 

```{r}
bac.prophylax <- mergeContainer %>% filter(!is.na(AntibacName.NeBPr.NeBAn)) %>% mutate(AntibacName.NeBPr.NeBAn = case_when(AntibacName.NeBPr.NeBAn == "Other antibacterials" & !is.na(otherAntibacText.NeBPr.NeBAn) ~ otherAntibacText.NeBPr.NeBAn, T ~ AntibacName.NeBPr.NeBAn ))

table(bac.prophylax$AntibacName.NeBPr.NeBAn)

 bac.prophylax %>% group_by(PatientID) %>% summarise(number.of.prophylax = n_distinct(AntibacName.NeBPr.NeBAn)) %>% pull(number.of.prophylax) %>% table()


```



### Distance to Samples for bacterial prophylax 

```{r}

bac.match <- mergeContainer %>% select(PatientID, AntibacName.NeBPr.NeBAn, AntibacDate.NeBPr.NeBAn,routeOfAdminstrationAntibac.NeBPr.NeBAn) %>% full_join(SampleDates, by= "PatientID") %>% filter(!is.na(AntibacDate.NeBPr.NeBAn) & PatientID %in% patientIDs ) #AntifungName.NeBPr.NeBAn, AntifungDate.NeBPr.NeBAn, AntiviralName.NeBPr.NeBAn, AntiviralProDate.NeBPr.NeBAn

## minimum distance between sample date and admission date of drug 

bac.match<- bac.match  %>% filter(as.Date(SampleDate) - as.Date(AntibacDate.NeBPr.NeBAn, format="%d.%m.%Y") >= 0) %>% 
  mutate( distance= as.Date(SampleDate) - as.Date(AntibacDate.NeBPr.NeBAn, format="%d.%m.%Y") )%>% 
  group_by(PatientID, SampleID) %>% slice_min(distance) %>% distinct()

summary(as.numeric(bac.match$distance))
hist(as.numeric(bac.match$distance))

n_distinct(bac.match$PatientID)

bac.match <- right_join(select(basic.info, SampleID=name, PatientID, Transplantation_Date, Event_Date), bac.match, by=c("PatientID","SampleID")) 

summary(as.numeric(as.Date(bac.match$AntibacDate.NeBPr.NeBAn, format="%d.%m.%Y") - bac.match$Transplantation_Date))


bac.prophylax_timeline <- mergeContainer %>% select(PatientID, AntibacName.NeBPr.NeBAn, AntibacDate.NeBPr.NeBAn,routeOfAdminstrationAntibac.NeBPr.NeBAn,otherAntibacText.NeBPr.NeBAn)  %>% 
mutate(AntibacName.NeBPr.NeBAn = case_when(AntibacName.NeBPr.NeBAn == "Other antibacterials" & !is.na(otherAntibacText.NeBPr.NeBAn) ~ otherAntibacText.NeBPr.NeBAn, 
                                           T ~ AntibacName.NeBPr.NeBAn )) %>%
  filter(!is.na(AntibacDate.NeBPr.NeBAn) & PatientID %in% patientIDs ) %>% 
  left_join(distinct(select(basic.info, PatientID, Transplantation_Date, Event_Date)), by="PatientID") %>% 
  relocate(Transplantation_Date, Event_Date, .after=PatientID) %>% relocate(AntibacName.NeBPr.NeBAn, .before=routeOfAdminstrationAntibac.NeBPr.NeBAn) %>% 
  distinct() %>% 
  mutate(AntibacDate.NeBPr.NeBAn = as.Date(AntibacDate.NeBPr.NeBAn,format="%d.%m.%Y")) %>%
  mutate(Tx_distance=AntibacDate.NeBPr.NeBAn - Transplantation_Date) %>%
    select(-otherAntibacText.NeBPr.NeBAn)

ggplot(bac.prophylax_timeline, aes(x=as.numeric(Tx_distance))) + geom_histogram()



```
##1.2 Viral Prophylax 


```{r}
viral.match <- mergeContainer %>% select(PatientID, AntiviralName.NeBPr.NeBAn, AntiviralProDate.NeBPr.NeBAn,routeOfAdminstrationAntiviral.NeBPr.NeBAn) %>% full_join(SampleDates, by= "PatientID") #AntifungName.NeBPr.NeBAn, AntifungDate.NeBPr.NeBAn, AntiviralName.NeBPr.NeBAn, AntiviralProDate.NeBPr.NeBAn


viral.match<- viral.match  %>% filter(as.Date(SampleDate) - as.Date(viral.match$AntiviralProDate.NeBPr.NeBAn, format="%d.%m.%Y") >= 0) %>% 
  mutate( distance= as.Date(SampleDate) - as.Date(AntiviralProDate.NeBPr.NeBAn, format="%d.%m.%Y") )%>% 
  group_by(PatientID, SampleID) %>% slice_min(distance) %>% distinct()


viral.prophylax_timeline <- mergeContainer %>% select(PatientID, AntiviralProDate.NeBPr.NeBAn, AntiviralName.NeBPr.NeBAn, routeOfAdminstrationAntiviral.NeBPr.NeBAn)  %>% 
  filter(!is.na(AntiviralProDate.NeBPr.NeBAn) & PatientID %in% patientIDs ) %>% 
  left_join(distinct(select(basic.info, PatientID, Transplantation_Date, Event_Date)), by="PatientID") %>% 
  relocate(Transplantation_Date, Event_Date, .after=PatientID) %>%
  distinct() %>% 
  mutate(AntiviralProDate.NeBPr.NeBAn = as.Date(AntiviralProDate.NeBPr.NeBAn,format="%d.%m.%Y")) %>%
  mutate(Tx_distance=AntiviralProDate.NeBPr.NeBAn - Transplantation_Date)

summary(as.numeric(viral.prophylax_timeline$Tx_distance))



summary(as.numeric(viral.match$distance))
table(viral.match$AntiviralName.NeBPr.NeBAn)

n_distinct(viral.match$PatientID)

```



##1.3 Fungal Prophylax 

```{r}
fungal.match <- mergeContainer %>% select(PatientID, AntifungName.NeBPr.NeBAn, AntifungDate.NeBPr.NeBAn,routeOfAdminstrationAntifung.NeBPr.NeBAn) %>% full_join(SampleDates, by= "PatientID") #AntifungName.NeBPr.NeBAn, AntifungDate.NeBPr.NeBAn, AntiviralName.NeBPr.NeBAn, AntiviralProDate.NeBPr.NeBAn


fungal.match<- fungal.match  %>% filter(as.Date(SampleDate) - as.Date(AntifungDate.NeBPr.NeBAn, format="%d.%m.%Y") >= 0) %>% 
  mutate( fung.distance= as.Date(SampleDate) - as.Date(AntifungDate.NeBPr.NeBAn, format="%d.%m.%Y") )%>% 
  group_by(PatientID, SampleID) %>% slice_min(fung.distance) %>% distinct()

summary(as.numeric(fungal.match$fung.distance))
table(fungal.match$AntifungName.NeBPr.NeBAn)

n_distinct(fungal.match$PatientID)

fungal.prophylax_timeline <- mergeContainer %>% select(PatientID,AntifungDate.NeBPr.NeBAn,  AntifungName.NeBPr.NeBAn, routeOfAdminstrationAntifung.NeBPr.NeBAn)  %>% 
  filter(!is.na(AntifungDate.NeBPr.NeBAn) & PatientID %in% patientIDs ) %>% 
  left_join(distinct(select(basic.info, PatientID, Transplantation_Date, Event_Date)), by="PatientID") %>% 
  relocate(Transplantation_Date, Event_Date, .after=PatientID) %>%
  distinct() %>% 
  mutate(AntifungDate.NeBPr.NeBAn = as.Date(AntifungDate.NeBPr.NeBAn,format="%d.%m.%Y")) %>%
  mutate(Tx_distance=AntifungDate.NeBPr.NeBAn - Transplantation_Date)


```
##1.4 Merging Prophylaxis 

```{r}
prophylaxis_timeline <- full_join(full_join(fungal.prophylax_timeline, viral.prophylax_timeline, by=c("PatientID", "Transplantation_Date", "Event_Date")), bac.prophylax_timeline, by=c("PatientID", "Transplantation_Date", "Event_Date"))


fungal.prophylax_timeline<- fungal.prophylax_timeline %>% rename(Prophylaxis_Date=AntifungDate.NeBPr.NeBAn, Prophylaxis_Name= AntifungName.NeBPr.NeBAn, Prophylaxis_Adminstration=routeOfAdminstrationAntifung.NeBPr.NeBAn) %>% mutate(Type= rep("fungal", nrow(fungal.prophylax_timeline)))

viral.prophylax_timeline <-  viral.prophylax_timeline  %>% rename(Prophylaxis_Date=AntiviralProDate.NeBPr.NeBAn, Prophylaxis_Name= AntiviralName.NeBPr.NeBAn, Prophylaxis_Adminstration=routeOfAdminstrationAntiviral.NeBPr.NeBAn) %>% mutate(Type= rep("viral", nrow(viral.prophylax_timeline)))


bac.prophylax_timeline <- bac.prophylax_timeline %>% rename(Prophylaxis_Date=AntibacDate.NeBPr.NeBAn, Prophylaxis_Name= AntibacName.NeBPr.NeBAn, Prophylaxis_Adminstration=routeOfAdminstrationAntibac.NeBPr.NeBAn) %>% mutate(Type= rep("bacterial", nrow(bac.prophylax_timeline)))


prophylaxis_timeline <- rbind(fungal.prophylax_timeline, bac.prophylax_timeline, viral.prophylax_timeline ) %>% arrange(PatientID) %>% relocate(Type, .after=Event_Date)


write.table(prophylaxis_timeline, "MetaDataProcessing/prophylaxis_timeline.tsv", sep="\t", quote = F, row.names = F, col.names = T)

paste(unique(prophylaxis_timeline$Prophylaxis_Name), collapse=".")


write.table(prophylaxis_timeline %>% filter(PatientID %in% cohort2_patients), "../tables/ClinicalData/Cohort2/prophylaxis_timeline.tsv", sep="\t", quote = F, row.names = F, col.names = T)

```

##1.5 Modification of the Prophylaxis Table 

some entries of prophylaxis are no prophylaxis but full on antimicrobial treatment
-> deleting them from the table 

```{r}
## wrong date given, correction:
prophylaxis_timeline[prophylaxis_timeline$PatientID == "Y980VYP6", "Tx_distance"] <- 1

## based on Felix B. 
not_prophylaxis <- prophy_timeline %>% filter(Prophylaxis_Type == "bacterial" & ((PatientID == "2EJGX6A4" & Tx_distance >=14) | 
                             (PatientID == "58T4MHL7" & Tx_distance %in% c(5,6,14,24)) | 
                              (PatientID == "EJWTQU66" & Tx_distance %in% c(11,13)) | 
                              (PatientID == "JEG9X7C9" & Tx_distance %in% c(22,13)) | 
                              (PatientID == "VU6NR3G2" & Tx_distance %in% c(13)))
                           )

not_prophylaxis <- not_prophylaxis %>% select(Container,Container_Date, PatientID,  AntibacDate=Prophylaxis_Date, AntibacName = Prophylaxis_Name, routeOfAdminstrationAntibac= Prophylaxis_Adminstration )
  

prophy_timeline <- prophy_timeline %>%filter(!(Prophylaxis_Type == "bacterial" & ((PatientID == "2EJGX6A4" & Tx_distance >=14) | 
                             (PatientID == "58T4MHL7" & Tx_distance %in% c(5,6,14,24)) | 
                              (PatientID == "EJWTQU66" & Tx_distance %in% c(11,13)) | 
                              (PatientID == "JEG9X7C9" & Tx_distance %in% c(22,13)) | 
                              (PatientID == "VU6NR3G2" & Tx_distance %in% c(13)))
                           ))

not_prophylaxis$antibiotic.class <- c("cephalosporin", "carbapenem", "linezolid", "cephalosporin","linezolid","carbapenem","linezolid","cephalosporin","fluorquinolone", "fosfomycin", "carbapenem", "ureidopenicillin", "cephalosporin" )

not_prophylaxis <- not_prophylaxis %>% left_join(basic.info %>% select(PatientID, Transplantation_Date, Event_Date, rej) %>% distinct(), by="PatientID")
```

##1.6 Prophylaxis Statistics for C2

```{r}
### complete C2
c2_prophy_timeline <- prophy_timeline %>% filter(PatientID %in% cohort2_patients) %>% mutate(Prophylaxis_Date = case_when(PatientID == "Y980VYP6" ~ as.Date("2017-09-22"), T ~ Prophylaxis_Date), Tx_distance= Prophylaxis_Date-Transplantation_Date)

c2_prophy_timeline %>% group_by(PatientID,rej) %>% filter(abs(Tx_distance) <90)  %>% summarise(type_prophy= n_distinct(Prophylaxis_Type), bac=any(Prophylaxis_Type=="bacterial"), viral=any(Prophylaxis_Type=="viral"), fungal=any(Prophylaxis_Type=="fungal") ) %>% group_by(rej) %>% summarise(n=n(), num_bac=length(bac[bac=TRUE]),                                                                                                                   per_bac=length(bac[bac=TRUE])/n(),
                num_viral=length(viral[viral==TRUE]),                                                                                                                                                                                                                                                                                 per_viral=length(viral[viral==TRUE])/n(),num_fungal=length(fungal[fungal==TRUE]),
                per_fungal=length(fungal[fungal==TRUE])/n())
```


```{r}
### prophylaxis overview n=92 c2: 

reducedc2_prophy_timeline <- prophy_timeline %>% filter(PatientID %in% in_patients) %>% mutate(Prophylaxis_Date = case_when(PatientID == "Y980VYP6" ~ as.Date("2017-09-22"), T ~ Prophylaxis_Date), Tx_distance= Prophylaxis_Date-Transplantation_Date)


write.table(reducedc2_prophy_timeline, "../tables/ClinicalData/Cohort2/C2_n92_prophylaxis_timeline.tsv", sep="\t", quote = F, row.names = F, col.names = T)

# %>% filter(abs(Tx_distance) <90)

reducedc2_prophy_timeline %>% group_by(PatientID,rej)   %>% summarise(type_prophy= n_distinct(Prophylaxis_Type), bac=any(Prophylaxis_Type=="bacterial"), viral=any(Prophylaxis_Type=="viral"), fungal=any(Prophylaxis_Type=="fungal") ) %>% group_by(rej) %>% summarise(n=n(), num_bac=length(bac[bac=TRUE]),                                                                                                                   per_bac=length(bac[bac=TRUE])/n(),
                num_viral=length(viral[viral==TRUE]),                                                                                                                                                                                                                                                                                 per_viral=length(viral[viral==TRUE])/n(),num_fungal=length(fungal[fungal==TRUE]),
                per_fungal=length(fungal[fungal==TRUE])/n())


```

## 1.7 Initial Prophylaxis C2 Plot 
based on reduced_prophy_timeline, prophlyaxis binned by Felix B. 
```{r}
initial_prophy<- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Rosa/tables/20240422_prophylaxis.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) 

addline_format <- function(x,...){
    gsub(' ','\n',  gsub('\\)','\\)\n',  gsub('\\+ ','\\+\n', gsub('/','\n',x))))
}


initial_prophy %>% group_by(rej) %>% filter(!is.na(cotrimoxazole)) %>% mutate(n=n_distinct(PatientID)) %>% group_by(rej,n) %>% summarise_at(vars(-PatientID), ~ sum(. == "yes")) %>% gather( "drug", "counts",-rej,-n) %>% mutate( perc=case_when(rej==0 ~ counts/n, T ~ counts/n), rej=case_when(rej==0 ~ "Normal Progess", T ~ "Rejection"))%>% ggplot(aes(x=reorder(drug, -counts), y=perc, fill=rej)) + geom_bar(stat="identity", position="dodge") + geom_text(aes(label=counts),position = position_dodge(0.9),vjust = 1.5,size=2.8) +  scale_fill_manual(na.translate = F, labels=c("Non Rejection", "Rejection"), values=c( "#0072B5FF","#BC3C29FF")) + theme_classic() + 
   theme(
    axis.line = element_line(colour = "darkgrey"),
    axis.ticks.y  = element_line(color = "black"),
    axis.text.y = element_text(color="black", size=8),
    
    axis.text.x = element_text(color="black", size=8),
    axis.ticks.x = element_line(color = "black"),
    
    axis.title.x  =element_text(size=13),
    axis.title.y  =element_text(size=13),

    legend.title = element_blank(),
    legend.text =  element_text(color="black", size=8),
    legend.position = "top",

    panel.border = element_blank(),
    panel.background = element_blank()
    ) + xlab("") + ylab("") +  scale_x_discrete(labels=addline_format(names(initial_prophy)[-c(1,2)]))


ggsave("../plots/cohort2/description/distribution_prophylax.pdf", width=8.5, height=4.3)

```


#2. Infections and Complications 

##2.1 Bacterial Infections

###2.1.1 Reading Files
1.4.1 Bact-NeBAntibacterials -> infections "BL.Ba.NeBBa"
2.3.1 Bact-NeFAntibacterials -> complications and antibactrial treatment, separated in two files

```{r}
## 1.4.1  Infection File 
bacterial_infections <- mergeContainer %>% select(PatientID, Container, ends_with("BL.Ba.NeBBa"),Container_Date=Datum..Container..x.x.x.x.x.x.x.x) %>% 
                                            mutate(AntibacName.BL.Ba.NeBBa= case_when(
  (AntibacName.BL.Ba.NeBBa == "Other antibacterials") & ! is.na(otherAntibacText.BL.Ba.NeBBa) ~ otherAntibacText.BL.Ba.NeBBa,
  T ~ AntibacName.BL.Ba.NeBBa
)) %>% filter(PatientID %in% patientIDs)  #pull(AntiBacName) %>% table()

bacterial_infections <- bacterial_infections %>% filter(rowSums(is.na(bacterial_infections)) != ncol(bacterial_infections)-3)

colnames(bacterial_infections)
```


```{r}
 #2.3.1Antibacterial Treatment File

X2_3_1Bact_NeFAntibacterials_filtered <- read_delim("Oct22/all_files_containing_patient_information/filtered/2.3.1Bact-NeFAntibacterials_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

X2_3_1Bact_NeFAntibacterials_filtered  %>% mutate(AntiBacName= case_when(
  (`AntibacName-FU-Ba-NeFBa` == "Other antibacterials") & ! is.na(`otherAntibacText-FU-Ba-NeFBa`) ~ `otherAntibacText-FU-Ba-NeFBa`,
  T ~ `AntibacName-FU-Ba-NeFBa`
)) %>% filter(PatientID %in% patientIDs) %>% pull(AntiBacName) %>% table()

#2.3.1 Bacterial Complication File

X2_3_1Bacterialcomplications_filtered <- X2_3_1Bacterialcomplications_filtered %>% filter(!is.na(Container) & PatientID %in% patientIDs) 

## 121 with complication entries 
X2_3_1Bacterialcomplications_filtered %>%  filter(! is.na(`diagDate-bacInfect-FU-Ba-NeFBa`) ) %>%  summarise(n_distinct(PatientID))


lapply(select(X2_3_1Bacterialcomplications_filtered,  ! contains ("#") ), table)

lapply(select(X2_3_1Bacterialcomplications_filtered,  ! contains ("Surgical") | contains("kidney") | contains("surgery") ), table)


table(X2_3_1Bacterialcomplications_filtered$`primaryOrReactivation-bacInfect-FU-Ba-NeFBa`)


bacterial_complications<- full_join(X2_3_1Bact_NeFAntibacterials_filtered, X2_3_1Bacterialcomplications_filtered, by=intersect(colnames(X2_3_1Bact_NeFAntibacterials_filtered) , colnames(X2_3_1Bacterialcomplications_filtered))
) %>%rename(Container_Date=`Datum (Container)`)
```


```{r}
## merge preparation

colnames(bacterial_complications) <- str_replace(colnames(bacterial_complications), "-FU-Ba-NeFBa", "")

colnames(bacterial_complications) <- str_replace(colnames(bacterial_complications), "-", ".")


bacterial_complications <- bacterial_complications[,-c(2,4,6:8)]

colnames(bacterial_infections) <- str_replace(colnames(bacterial_infections), ".BL.Ba.NeBBa", "")

intersect(colnames(bacterial_complications) ,colnames(bacterial_infections)) 


## merging together 

bac_infec_compl<- full_join(bacterial_complications, bacterial_infections, by= intersect(colnames(bacterial_complications) ,colnames(bacterial_infections)) 
) %>% arrange(PatientID) %>% distinct()
```


```{r}
### trying to gather all site columns to one site column 
site_info <- bac_infec_compl %>% gather("site_bacInfect", "value",  starts_with("site.bacInfect") | starts_with("otherSite"), na.rm = T) 

site_info <- site_info %>% mutate(site_bacInfect = case_when(
  startsWith(site_bacInfect, "other") ~ value,
  endsWith("Sitenotidentified", site_bacInfect ) ~ NA,
  T~ sub(".*#", "", site_bacInfect  )
)) %>% select( -value)


na_site_info <- bac_infec_compl %>%
  filter(rowSums(is.na(select(., starts_with("site.bacInfect") | starts_with("otherSite") ))) == ncol(select(., starts_with("site.bacInfect") | starts_with("otherSite"))))

na_site_info <- na_site_info %>% mutate(site_bacInfect=rep(NA, nrow(na_site_info))) %>% select(-starts_with("site.bacInfect"), -starts_with("otherSite") )
# Now filtered_df contains only those rows where all "site-bacInfect" columns have NA values

bac_infec_compl <- rbind(site_info, na_site_info) %>% arrange(PatientID) %>% 
  group_by_at(vars(-site_bacInfect)) %>% summarise(site_bacInfect= paste(site_bacInfect, collapse=";")) %>% ungroup() 

bac_infec_compl$site_bacInfect
```


```{r}
## adding miss classified prophlyaxis to bact.infection.compl. table , see Section 1.5 
bac_infec_compl<- bac_infec_compl %>% mutate(Container_Date = as.Date(Container_Date, format= "%d.%m.%Y") , AntibacDate = as.Date(AntibacDate, format= "%d.%m.%Y"))%>% full_join(not_prophylaxis,., by= names(not_prophylaxis))


```

```{r}
#adding Tx Event Dates and rej status 
bac_infections <- right_join(basic.info %>% select(PatientID, Transplantation_Date, Event_Date,rej) %>% distinct(),bac_infec_compl %>% select(-contains("AGEFor")), by="PatientID")  



bac_infections <- bac_infections %>% relocate(Container, Container_Date) %>% 
                           relocate(diagDate_bacInfect=diagDate.bacInfect, Antibac_Date=AntibacDate, .after=rej) %>%
                           mutate_at(vars(diagDate_bacInfect, Container_Date, Antibac_Date) , as.Date, format="%d.%m.%Y")

```

###2.1.2 Adding Antibiotic Classes

defined by Felix B. for C2
15 antibiotic classes: 
"fluorquinolone",
"cephalosporin",
"carbapenem",
"ureidopenicillin",
"glycopeptide",
"aminopenicillin",
"trimethoprim/sulfomethoxazole",
"penicillin"                   
"other",
"macrolide",
"amidinopenicillin",            
"fosfomycin",
"linezolid",
"clindamycin",                  
"metronidazole"

```{r}
bac_infectionsFB <- read.csv2("../tables/bacterial_infection_FB.tsv", sep="\t") %>% mutate_at(vars(contains("Date")), funs(as.Date(.))) %>% mutate(rej=as.factor(rej))

all(names(bac_infections) %in% intersect(names(bac_infections), names(bac_infectionsFB)) )

bac_infections <-  bac_infections %>% filter(!(PatientID %in% cohort2_patients)) %>% 
  full_join(bac_infectionsFB,., by=intersect(names(bac_infections), names(bac_infectionsFB))) %>% 
  full_join(not_prophylaxis, by=intersect(names(not_prophylaxis), names(.))) %>% mutate(antibiotic.class =case_when(antibiotic.class=="" ~ NA, T ~antibiotic.class))
```




###2.1.3 Redefining Bac. Infections

primaryOrReactivation.bacInfect and OLDclinTypeBacInfectHIDDEN providing information about infection type
suspected, proven, primary infection,  possible disease, probable disease, and colonization 
sorting out 
```{r}
#bac_infections %>% select(primaryOrReactivation.bacInfect, "typeVerfiedBy.Medicalreport" ,  "typeVerfiedBy.Consultationofattendingdoctor", "OLDclinTypeBacInfectHIDDEN") %>% filter(OLDclinTypeBacInfectHIDDEN == "Colonization") %>% apply(.,2, table)

## replace nas, with old type
transformed_bac_infections <- bac_infections %>% mutate(primaryOrReactivation.bacInfect = case_when(is.na(primaryOrReactivation.bacInfect) | primaryOrReactivation.bacInfect == "" ~ OLDclinTypeBacInfectHIDDEN, T~primaryOrReactivation.bacInfect))

table(transformed_bac_infections$primaryOrReactivation.bacInfect)

## remove suspected infections 
transformed_bac_infections<- transformed_bac_infections %>% filter(!primaryOrReactivation.bacInfect %in% c("Possible disease", "Probable disease", "suspected infection") | is.na(primaryOrReactivation.bacInfect)) 

table(transformed_bac_infections$primaryOrReactivation.bacInfect)

# either antibac admin or reactivation type or antibiotic treatment

transformed_bac_infections<- transformed_bac_infections %>% mutate(AntibacName = na_if(AntibacName, ""), primaryOrReactivation.bacInfect=na_if(primaryOrReactivation.bacInfect,"")) 

transformed_bac_infections %>% filter(is.na(primaryOrReactivation.bacInfect))


transformed_bac_infections<- transformed_bac_infections %>% filter(!is.na(AntibacName) | !is.na(primaryOrReactivation.bacInfect) | antibioticTreatment == "yes" )

## if diag date is not given, take antibac date
summary(transformed_bac_infections$diagDate_bacInfect)

transformed_bac_infections<- transformed_bac_infections %>% mutate(diagDate_bacInfect = case_when(is.na(diagDate_bacInfect) ~ Antibac_Date, T ~ diagDate_bacInfect))

summary(transformed_bac_infections$diagDate_bacInfect)

## check whether entries are in close proximity and remove them if its 5 days or less
transformed_bac_infections<- transformed_bac_infections %>% group_by(PatientID) %>% arrange(PatientID, diagDate_bacInfect) %>%
  mutate(
    # Calculate the difference between current and previous diagnosis dates
    diff_to_next_infection = c(NA, difftime(diagDate_bacInfect[-1], diagDate_bacInfect[-n()], units = "days")),
    # Flag rows where the difference is less than or equal to 5 days
    flag = ifelse(diff_to_next_infection <= 5, TRUE, FALSE),
  
  ) %>% ungroup() 


table(transformed_bac_infections$diff_to_next_infection)
summary(transformed_bac_infections$diff_to_next_infection)

## for one infection several entries, due to differet given antibiotics or other reason, keeping them in for the antibiotic variables 

```


### 2.1.4 Plot Antibiotic Usage C2

```{r}
bac_infection_cohort2 <-  transformed_bac_infections %>% filter(PatientID %in% in_patients)  %>%   mutate(Antibac_Date= case_when(is.na(Antibac_Date) &  !is.na(diagDate_bacInfect) ~ diagDate_bacInfect, T ~ Antibac_Date )) # %>% filter(! is.na(Antibac_Date) | ! is.na(AntibacName))
  
  
bac_infection_cohort2<- bac_infection_cohort2 %>% mutate(Tx_AntiBac_distance= Antibac_Date - Transplantation_Date, 
                                                         Event_AntiBac_distance= Antibac_Date - Event_Date,
                                                         .after=rej )  %>% filter(!is.na(antibiotic.class) & antibiotic.class !="")

bac_infection_cohort2 %>% group_by(PatientID,rej) %>% summarise(num_different_antibac = n_distinct(antibiotic.class,na.rm = T) , num_antibac= n()) %>% mutate(num_antibac = case_when(num_different_antibac ==0 ~ 0, T ~ num_antibac)) %>%
  ggplot(aes(x=rej, y= num_antibac, fill=rej)) + geom_boxplot() +
  stat_compare_means(label.x.npc = "center", size=5) + xlab("") + ylab("Number of Antibacterials") +theme(legend.position = "none") + scale_x_discrete(label= c("Normal Progress", "Rejection"))



all_132_cases <-bac_infection_cohort2 %>% mutate(grp= case_when(rej==0 ~ "NormalProgress", T ~ "BeforeRejection")) %>% rbind(bac_infection_cohort2 %>% filter(rej==1) %>% mutate(grp= "AfterRejection")) %>% rbind(bac_infection_cohort2 %>% filter(rej==1) %>% mutate(grp= "Rejection")) %>% relocate(grp,.after=Event_AntiBac_distance ) %>% distinct(PatientID, grp)



plot_infection_data <- bac_infection_cohort2 %>% mutate(grp= case_when(rej==0 ~ "NormalProgress", T ~ "BeforeRejection")) %>% rbind(bac_infection_cohort2 %>% filter(rej==1) %>% mutate(grp= "AfterRejection"))%>% rbind(bac_infection_cohort2 %>% filter(rej==1) %>% mutate(grp= "Rejection"))%>% relocate(grp,.after=Event_AntiBac_distance ) %>% filter((grp=="AfterRejection" & Event_AntiBac_distance >0 & !is.na(Event_AntiBac_distance)) |
                                          (grp=="BeforeRejection" & Event_AntiBac_distance <0 & !is.na(Event_AntiBac_distance)) | 
                                            (grp=="Rejection"  & !is.na(Event_AntiBac_distance)) |
                                            (grp=="NormalProgress" & is.na(Event_AntiBac_distance)) )%>%
  
   #mutate(antibiotic.class = case_when(antibiotic.class== "" ~ NA, T~ antibiotic.class)  )  %>% 
 # distinct(PatientID, Tx_AntiBac_distance, Event_AntiBac_distance, rej, grp, Antibac_Date, antibiotic.class,routeOfAdminstrationAntibac) %>% 
  group_by(PatientID,grp) %>%
  summarise(num_different_antibac = n_distinct(antibiotic.class[Tx_AntiBac_distance >= 0  & !is.na(Antibac_Date)],na.rm = T) , 
            num_antibac= length(antibiotic.class[Tx_AntiBac_distance >= 0  & !is.na(Antibac_Date) ]), 
            num_parenteral= sum(routeOfAdminstrationAntibac[Tx_AntiBac_distance >= 0 & !is.na(Antibac_Date)] == "parenteral")) %>% 
  
  right_join(.,all_132_cases, by=c("PatientID", "grp")) %>% replace(is.na(.),0) %>% 
  
  gather("type", "number", num_antibac,num_different_antibac, num_parenteral)

plot_infection_data$grp <- factor(plot_infection_data$grp,level=c("NormalProgress", "Rejection", "BeforeRejection", "AfterRejection"))
plot_infection_data$inter <- interaction( plot_infection_data$grp,plot_infection_data$type)

levels=levels(plot_infection_data$inter)
my_comparisons <- list(c(levels[1], levels[2]),
                       c(levels[3], levels[4]),
                       c(levels[5], levels[6]),
                       c(levels[7], levels[8]),
                       c(levels[9], levels[10]),
                       c(levels[11], levels[12])
                       )

#label.y=max(plot_infection_data$number)*1.03
my_comparisons <- list(c("NormalProgress", "Rejection"), c("BeforeRejection", "AfterRejection"))

antbiotic_admin_plot <-   ggplot(plot_infection_data,aes(x=grp,  y= number, fill=factor(grp,level=c("NormalProgress", "Rejection", "BeforeRejection", "AfterRejection"))))  + geom_boxplot() + facet_wrap(~type, labeller = labeller(type=c("num_antibac"="Total number of antibiotics\ntaken after KT", "num_different_antibac"="Number of different antibiotics\ntaken after KT", "num_parenteral"="Total number of\nparenteral administration")))+
 xlab("") + ylab("")  +   theme(legend.title = element_blank(),axis.ticks.x = element_blank(), axis.text.x = element_text(size=13), legend.text=element_text(size=13)) +  
    stat_compare_means(comparisons = my_comparisons, method = "wilcox.test", label = "p.format", label.y=max(plot_infection_data$number)*1.01) +
theme_bw() + 
  theme(
   strip.text = element_text(size=11),
    strip.background = element_blank(),
    strip.placement = "outside", # Remove facet borders
    line = element_line(colour = "black"),
    axis.ticks.y  = element_line(color = "black"),
    axis.text.y = element_text(color="black", size=9),
    axis.text.x = element_text(color="black", size=10),
    axis.ticks.x = element_line(color = "black"),
    axis.title.x  =element_text(size=10),
    axis.title.y  =element_text(size=10),
    legend.title = element_blank(),
    legend.position="none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) + scale_x_discrete(labels=c("Normal\nProgress", "Rejection", "Before\nRejection", "After\nRejection")) + scale_fill_manual(values=c( "#0072B5FF","#BC3C29FF","#E18727FF", "#20854EFF"))

  
ggsave("../plots/cohort2/description/removed_antibiotic_administration.pdf", antbiotic_admin_plot, width=10, height=6)

```



##2.2 Viral Infections

###2.2.1 Reading Files

2.3.2 for complications
1.4.2 for infections 


#### Infection File 

taken infections as a subset from the merged container indexed file:

```{r}
## 1.4.2 infectionsAV
select(mergeContainer, PatientID, AntiviralName.BL.Vi.NeBvi , otherAntiviralsText.BL.Vi.NeBvi)  %>% mutate(AntiViralName= case_when(
  (AntiviralName.BL.Vi.NeBvi == "Other antivirals") & ! is.na( otherAntiviralsText.BL.Vi.NeBvi) ~ otherAntiviralsText.BL.Vi.NeBvi,
  T ~AntiviralName.BL.Vi.NeBvi
)) %>% filter(PatientID %in% patientIDs) %>% distinct(PatientID, AntiViralName) %>% pull(AntiViralName) %>% table()
```


```{r}
viral_infections <- select(mergeContainer, PatientID, Container, Container_Date=Datum..Container..y.y.y.y.y.y.y.y,  contains("BL.Vi.NeBvi"))

colnames(viral_infections)<- sub(".BL.Vi.NeBvi", "", colnames(viral_infections))

table(viral_infections$AntiviralName)

viral_infections <- viral_infections %>%  mutate(AntiviralName = case_when(is.na(AntiviralName) ~ otherAntiviralsText, 
                                   AntiviralName == "Other antivirals"  ~ otherAntiviralsText, 
                                   T ~  AntiviralName)) %>% select(-otherAntiviralsText)

table(viral_infections$AntiviralName)
```


#### Complication Files

some general checking
```{r}
## 2.3 Complications 
X2_3_NeFVir_Viral_complicaAV_filtered <- read_delim("Oct22/all_files_containing_patient_information/filtered/2.3.-NeFVir_Viral_complicaAV_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

X2_3_NeFVir_Viral_complicaAV_filtered <- X2_3_NeFVir_Viral_complicaAV_filtered %>%  filter(PatientID %in% patientIDs)

X2_3_NeFVir_Viral_complicaAV_filtered  %>% mutate(AntiViralName= case_when(
  (`AntiviralName-FU-Vi-NeFVi` == "Other antivirals") & ! is.na( `otherAntiviralsText-FU-Vi-NeFVi`) ~`otherAntiviralsText-FU-Vi-NeFVi`,
  T ~ `AntiviralName-FU-Vi-NeFVi`
)) %>% distinct(PatientID, AntiViralName) %>% pull(AntiViralName)  %>% table()

```

```{r}
X2_3_2Viralcomplications_filtered <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Oct22/all_files_containing_patient_information/filtered/2.3.2Viralcomplications_filtered.csv",  delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% filter(!is.na(Container) & PatientID %in% patientIDs) 

X2_3_2Viralcomplications_filtered %>% filter(! is.na(`diagDate-virInfect-FU-Vi-NeFVi`) ) %>% summarise(n_distinct(PatientID))

table(X2_3_2Viralcomplications_filtered$`primaryOrReactivation-virInfect-FU-Vi-NeFVi`)

## postive testing for different viruses -> very sparse 
lapply(select(X2_3_2Viralcomplications_filtered,   contains ("kidney") ), table)


### belonging antiviral treatments 
X2_3_NeFVir_Viral_complicaAV_filtered <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle//Oct22/all_files_containing_patient_information/filtered/2.3.-NeFVir_Viral_complicaAV_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) %>% 
  filter(!is.na(Container) & PatientID %in% patientIDs)
```

###2.2.2 Merging Files
```{r}
#merging antivirals and complications together 

viral_complications <- full_join(X2_3_NeFVir_Viral_complicaAV_filtered, X2_3_2Viralcomplications_filtered, by=intersect(colnames(X2_3_2Viralcomplications_filtered) ,colnames(X2_3_NeFVir_Viral_complicaAV_filtered)))

colnames(viral_complications) <- sub("-", ".", sub("-FU-Vi-NeFVi", "",colnames(viral_complications) ))

## add other antiviral text to missing antiviral name 

viral_complications <- viral_complications %>% 
  mutate(AntiviralName = case_when(is.na(AntiviralName) ~ otherAntiviralsText, 
                                   AntiviralName == "Other antivirals"  ~ otherAntiviralsText, 
                                   T ~  AntiviralName)) %>% 
  select(-otherAntiviralsText, -Status, -Dokument,  -`Container Index`, -Erfassungsdatum, -`2.3..NeFVir_Viral_complicaAV`) %>% rename(Container_Date=`Datum (Container)`)



viral_infec_compl<- full_join(viral_infections,viral_complications, by= intersect(colnames(viral_infections), colnames(viral_complications))) %>%
                    mutate(diagDate_virInfect= as.Date(diagDate.virInfect, format="%d.%m.%Y"),
                           Antiviral_Date= as.Date(AntiviralProDate, format="%d.%m.%Y"),
                           Container_Date= as.Date(Container_Date, format="%d.%m.%Y"), .keep="unused", .after=Container) %>%
                    relocate(Container, Container_Date)

viral_infec_compl<- viral_infec_compl  %>% filter(rowSums(is.na(viral_infec_compl)) < ncol(viral_infec_compl)-2)


viral_infec_compl  %>% filter(is.na(diagDate_virInfect))
```

```{r}
vir_infections <- left_join( basic.info %>% select(PatientID, Transplantation_Date, Event_Date,rej) %>% distinct(),  viral_infec_compl , by="PatientID") %>% select( -contains("AGEf")) %>% 
                          relocate(Container, Container_Date) %>%
                          relocate(starts_with("fever"), .after=AntiviralName)


```


###2.2.3 Subsetting to interested Columns
```{r}
viral_subset <- vir_infections %>% select(Container, PatientID, rej, Transplantation_Date, Event_Date, diagDate_virInfect, cmvPCR, cmv2PCR,  bkvPCR,bkv2PCR,	ebvPCR, ebv2PCR,
primaryOrReactivation.virInfect, OLDclinTypeVirInfectHIDDEN) %>% filter(!is.na(Container)) %>% distinct()



## nas auffüllen von xx2PCR ? 

viral_subset<- viral_subset %>% mutate(cmvPCR = case_when(is.na(cmvPCR) & !is.na(cmv2PCR) ~ cmv2PCR, T ~ cmvPCR),
                       bkvPCR = case_when(is.na(bkvPCR) & !is.na(bkv2PCR) ~ bkv2PCR, T ~ bkvPCR) ,
                       ebvPCR = case_when(is.na(ebvPCR) & !is.na(ebv2PCR) ~ ebv2PCR, T ~ ebvPCR)) %>% select(-ends_with("2PCR"), -OLDclinTypeVirInfectHIDDEN, -Container)

viral_subset <- viral_subset %>% mutate(ViralInfec_Tx_Distance= diagDate_virInfect - Transplantation_Date, .after=diagDate_virInfect)
```




###2.2.4 Redefining Viral Infection

remove labeled suspected infection,  Detection of antibodies without signs of acute disease, not counting as viral infections
```{r}
## taken from vir_infections only diagnosis date and type of viral infections
transformed_viral_infec <- vir_infections %>% select(Container, PatientID, rej, Transplantation_Date, Event_Date, diagDate_virInfect,
primaryOrReactivation.virInfect, OLDclinTypeVirInfectHIDDEN) %>% filter(!is.na(diagDate_virInfect)) %>% distinct()

## adding distance to Tx
transformed_viral_infec <- transformed_viral_infec %>% mutate(ViralInfec_Tx_Distance= diagDate_virInfect - Transplantation_Date, .after=diagDate_virInfect)

## in case primary or reactivation is na, then take old clin type 
transformed_viral_infec %>% mutate(primaryOrReactivation.virInfect = case_when(is.na(primaryOrReactivation.virInfect) ~ OLDclinTypeVirInfectHIDDEN, T ~ primaryOrReactivation.virInfect)) 

## getting rid of all Detection of antibodies without signs of acute disease and suspected infection
transformed_viral_infec <- transformed_viral_infec %>% filter(primaryOrReactivation.virInfect != "suspected infection" & primaryOrReactivation.virInfect != "Detection of antibodies without signs of acute disease")

## check whether entries are in close proximity and remove them if its 5 days or less
transformed_viral_infec<- transformed_viral_infec %>% group_by(PatientID) %>% arrange(PatientID, diagDate_virInfect) %>%
  mutate(
    # Calculate the difference between current and previous diagnosis dates
    diff_to_next_infection = c(NA, difftime(diagDate_virInfect[-1], diagDate_virInfect[-n()], units = "days")),
    # Flag rows where the difference is less than or equal to 5 days
    flag = ifelse(diff_to_next_infection <= 5, TRUE, FALSE),
  
  ) %>% ungroup() %>%
## for two patients , flag==T, two entries on the same diag date is given -> remove one of them 
  filter(!flag |is.na(flag) )

summary(transformed_viral_infec$diff_to_next_infection)



```

###2.2.5 Plots of Viral Infection C2
```{r}
# everything after kt 
data.frame(basic.info %>% filter(PatientID %in% cohort2_patients) %>% select(PatientID, rej) %>% distinct()) %>% left_join(.,transformed_viral_infec, by=c("PatientID","rej") ) %>% filter(ViralInfec_Tx_Distance >=0 | is.na(ViralInfec_Tx_Distance)) %>% group_by(PatientID,rej) %>% summarise(n=n_distinct(diagDate_virInfect, na.rm = T)) %>%
  ggplot(aes(x=rej, y=n, fill=rej)) +  geom_boxplot() +  stat_compare_means(label.x.npc = "center", size=4 , method= "wilcox.test") + xlab("") + ylab("Number of Viral Infections after KT")  +theme(legend.position = "none") + scale_x_discrete(label= c("Normal Progress", "Rejection"))
```


```{r}
## time frame kt to sample

viral_infec_plot <- data.frame(basic.info %>% filter(PatientID %in% cohort2_patients) %>% select(PatientID, SampleID, Sample_Date=Date, rej) %>% distinct()) %>% left_join(.,transformed_viral_infec, by=c("PatientID","rej") ) %>% filter(ViralInfec_Tx_Distance >=0 | is.na(ViralInfec_Tx_Distance)) %>% group_by(PatientID,rej) %>%
  
  summarise(KT_Sample=n_distinct(diagDate_virInfect[diagDate_virInfect <= Sample_Date], na.rm = T), all_KT=n_distinct(diagDate_virInfect, na.rm = T)) %>%
  gather("time_frame", "value", KT_Sample,all_KT) %>%
  
  ggplot(aes(x=rej, y=value, fill=rej)) +  geom_boxplot() +  stat_compare_means(label.x.npc = "center", size=3.3, method= "wilcox.test", label.x=1.2) + xlab("") + ylab("Number of Viral Infections")  + facet_wrap(~time_frame, labeller = labeller(time_frame = 
    c("all_KT" = "all follow-ups after KT", "KT_Sample"= "time span KT to Sample"))) +  theme_bw() +
    theme(
   strip.text = element_text(size=12),
    strip.background = element_blank(),
    strip.placement = "outside", # Remove facet borders
    line = element_line(colour = "black"),
    axis.ticks.y  = element_line(color = "black"),
    axis.text.y = element_text(color="black", size=11),
    axis.text.x = element_text(color="black", size=12),
    axis.ticks.x = element_line(color = "black"),
    axis.title.x  =element_text(size=12),
    axis.title.y  =element_text(size=12),
    legend.title = element_blank(),
    legend.position="none",
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()
  ) + 
    scale_x_discrete(label= c("Normal Progress", "Rejection")) +  scale_fill_nejm()
  scale_fill_manual(breaks=c(0, 1), values=c( "#0072B5FF","#BC3C29FF")) 

ggsave("../plots/cohort2/description/latex_viral_infections.pdf", viral_infec_plot, width=7, height=4, dpi=300)
#+ scale_x_discrete(label= c("Normal Progress", "Rejection"))
```


#3. Fungal Infections

##3.1. Reading Files
- 1.4.3 only one patient with infection
-2.3.3 fungal complications 
-2.3 AF: belonging antifungal treatments 


```{r}

#### Complications 
X2_3_3Fungalcomplications_filtered <- read_delim("../../Oct22/all_files_containing_patient_information/filtered/2.3.3Fungalcomplications_filtered.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE)

X2_3_3Fungalcomplications_filtered <- X2_3_3Fungalcomplications_filtered  %>% filter(!is.na(Container) & PatientID %in% patientIDs) 

X2_3_3Fungalcomplications_filtered %>% filter(! is.na(`diagDate-fungInfect-FU-Fu-NeFMy`) ) %>% summarise(n_distinct(PatientID))

table(X2_3_3Fungalcomplications_filtered$`primaryOrReactivation-fungInfect-FU-Fu-NeFMy`)

 X2_3_3Fungalcomplications_filtered <- X2_3_3Fungalcomplications_filtered%>% select(-Dokument, -Erfassungsdatum, -`Container Index`, -Status, -contains("AGEF"))


colnames(X2_3_3Fungalcomplications_filtered) 

### belonging anti fungal treatments
X2_3_NeFMyc_Fungal_complicaAF <- read_delim("/fast/AG_Forslund/shared/DZIF_CKD_Holle/Oct22/all_files_containing_patient_information/2.3-NeFMyc_Fungal_complicaAF.csv", 
    delim = ";", escape_double = FALSE, trim_ws = TRUE) 

X2_3_NeFMyc_Fungal_complicaAF <- X2_3_NeFMyc_Fungal_complicaAF %>% select(-Dokument, -Erfassungsdatum, -`Container Index`, -Status, -`2.3-NeFMyc_Fungal_complicaAF`, -`otherAntifungText-FU-Fu-NeFMy`, -`AntifungDate-FU-Fu-NeFMyAGEForVariable`)


### add AF to complications 
X2_3_NeFMyc_Fungal_complicaAF$Container %in% X2_3_3Fungalcomplications_filtered$Container

fungal_complications<- right_join(X2_3_NeFMyc_Fungal_complicaAF, X2_3_3Fungalcomplications_filtered, by=c("PatientID","Container","Datum (Container)"))

## change column names 

colnames(fungal_complications) <- str_replace(colnames(fungal_complications), "-FU-Fu-NeFMy", "")
colnames(fungal_complications) <- str_replace_all(colnames(fungal_complications), "-", ".")

## change dates 

fungal_complications <- fungal_complications %>% relocate(Container, Container_Date=`Datum (Container)`) %>%
                                                 mutate(Container_Date= as.Date(Container_Date, format="%d.%m.%Y")) %>%
                                                 mutate(diagDate_fungInfect = as.Date(diagDate.fungInfect, format="%d.%m.%Y"),
                                                   Antifung_Date=  as.Date(AntifungDate, format="%d.%m.%Y"), 
                                                        .keep="unused", .after=PatientID) 
## add basic info 

fungal_complications <- basic.info %>% select(PatientID, Transplantation_Date, Event_Date, rej) %>% distinct() %>% right_join(fungal_complications, by="PatientID") %>% relocate(Container, Container_Date)



```

#4. Statistics for Cohort2 for all Infections

```{r}
## fungal infection 
selected_fungal_infec %>% remove_empty("cols")  %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% cohort2_Ids),., by="PatientID") %>% mutate(diff=diagDate_fungInfect - SampleDate) %>%  group_by(PatientID) %>% filter(diff < 0) %>% group_by(rej) %>% summarise(n())


## bac infection 
bac_infections %>% filter(PatientID %in% cohort2_patients)  %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% cohort2_Ids),., by="PatientID") %>%   mutate(diff=diagDate_bacInfect - SampleDate)  %>% filter(diff <=0 & !is.na(diff)) %>% distinct(SampleID, diagDate_bacInfect, rej, diff) %>% group_by(rej) %>% summarise(n=n(), patient=n_distinct(SampleID)) 

#table(bac_infections$primaryOrReactivation.bacInfect)


## transformed bac infection
## for not counting infection twice, diff_to_infection > 5
transformed_bac_infections %>% filter(PatientID %in% in_patients)  %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% in_samples),., by="PatientID") %>% filter(diff_to_next_infection > 5 | is.na(diff_to_next_infection)) %>%
mutate(diff=diagDate_bacInfect - SampleDate) %>% filter(diff <= 0 & !is.na(diff)) %>% distinct(SampleID,  rej, diff) %>% group_by(rej) %>% summarise(n=n(), patient=n_distinct(SampleID)) 

# viral infection

viral_subset %>% filter(PatientID %in% in_patients)  %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% cohort2_Ids),., by="PatientID") %>%   mutate(diff=diagDate_virInfect - SampleDate) %>% filter(diff <= 0 & !is.na(diff)) %>% distinct(SampleID,  rej, diff) %>% group_by(rej) %>% summarise(n=n(), patient=n_distinct(SampleID)) 

## transformed viral infection 

## numbers
transformed_viral_infec %>% filter(PatientID %in% in_patients)  %>% full_join(basic.info %>% select(SampleID, SampleDate=Date, PatientID) %>% filter(SampleID %in% in_samples),., by="PatientID") %>%   mutate(diff=diagDate_virInfect - SampleDate) %>% filter(diff <= 0 & !is.na(diff)) %>% distinct(SampleID,  rej, diff) %>% group_by(rej) %>% summarise(n=n(), patient=n_distinct(SampleID)) 

## types
transformed_viral_infec %>% filter(PatientID %in% in_patients & rej==1) %>% count(rej, primaryOrReactivation.virInfect)

viral_infec_compl
```

